---
title: "switchtube_log_analysis"
author: "Kai Lukoff"
date: "5/13/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE)
```

## Introduction

This RMarkdown shows the cleaning and analysis of the log data for the SwitchTube mobile app.

Log data is stored ion BigQuery in a SQL database from where it can be downloaded in JSON format. The JSON data is complex and nested and therefore first needs to be cleaned before analysis.

The desired data format is like this:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
desired_df <- tibble::tribble(
  ~event_id, ~param_device, ~action, ~param_video_id, ~Pets_Animals, ~etc,
  "247974748",   "Motorola E (4)",  "video_started",   "2B30zHE7DJ0",   "1",   "etc",
  "1745499870",   "Motorola E (4)",  "NA",   "NA",   "NA",   "etc",
  "850539252",   "Android Nexus VI",  "video_ended",   "9OIJOISF81",   "NA",   "etc",
  "etc", "etc",  "etc", "etc",  "etc", "etc",
)
desired_df
```

Libraries required for data cleaning:
```{r libraries, message=FALSE}
library("jsonlite")
library("tidyverse")
```


## Data import
Import the JSON data as a tibble
```{r import_json}

switchtube_tbl <- as_tibble(jsonlite::fromJSON("results-20210506-134120.json"))
```

Show data after import
```{r, warning=FALSE}
switchtube_tbl
```
Note that the event_params column contains nested dataframes that first need to be unnested to get at the right data.

# Data cleaning steps
1) Filter rows for only events that are study events
```{r}
switchtube_tbl2 <- switchtube_tbl %>% dplyr::filter(event_name == "study_event")
```

2) Select columns for only the event params column
```{r}
switchtube_tbl3 <- switchtube_tbl2 %>% select(event_params)
```

3) Flatten the data in the event_params column of the dataset (want just the study event data)
```{r}
switchtube_tbl4 <- as_tibble(unnest_longer(switchtube_tbl3,event_params))
```

4) Filter rows where event_params.key == study_event_data
```{r}
switchtube_tbl5 <- switchtube_tbl4 %>% dplyr::filter(event_params$key == "study_event_data")
```
Note: need to use dollar sign instead of period in the column name when filtering

5) Filter rows where event_params.key == study_event_data
```{r}
switchtube_tbl5 <- switchtube_tbl4 %>% dplyr::filter(event_params$key == "study_event_data")
```

6) Select columns for only the event params column
```{r}
switchtube_tbl6 <- as_tibble(switchtube_tbl5$event_params$value$string_value)
```
For some reason, this version returns an error: switchtube_tbl6 <- switchtube_tbl5 %>% select(event_params$value$string_value)

7) Add semi-colon to the end of each entry (should make it easier to deal with the merge later on)
```{r}
switchtube_tbl6$value <- paste0(switchtube_tbl6$value, ";")
```

8) Split columns at the semi-colon (into event_id and logs) using separate function in library(tidyr)
```{r}
switchtube_tbl7 <- switchtube_tbl6 %>% tidyr::separate(value, into = c("event_id", "logs"), sep = ";", extra="merge")
```
Source: https://stackoverflow.com/questions/55748363/how-to-split-a-dataframe-column-by-the-first-instance-of-a-character-in-its-valu

9) Group by event ID and unite results
```{r}
switchtube_tbl8 <- switchtube_tbl7 %>% group_by(event_id) %>% mutate(grouped_logs = paste0(logs, collapse = ""))
```

10) Select columns for only the event ID and united_results
```{r}
switchtube_tbl9 <- switchtube_tbl8 %>% select(event_id,grouped_logs)
```

11) De-duplicate rows
```{r}
switchtube_tbl10 <- switchtube_tbl9 %>% unique()
```

12) Remove final semi-colon from grouped-logs
```{r}
switchtube_tbl10$grouped_logs <- substr(switchtube_tbl10$grouped_logs,
                       start= 1, 
                       stop= nchar(switchtube_tbl10$grouped_logs)-1)
```

12) Remove id= from values in the event_id column
```{r}
switchtube_tbl10$event_id <- substr(switchtube_tbl10$event_id,
                       start= 4, 
                       stop= nchar(switchtube_tbl10$event_id))
```

## Current data
Show data after the above cleaning operations
```{r, warning=FALSE}
switchtube_tbl10
```

The grouped_logs column contains many key-value pairs separated by semi-colons. For example:
```{r, warning=FALSE}
switchtube_tbl10[1,2]
# not sure how to get this to print the entire string here rather than a shortened version
```
Here, the key is Pets + Animals and the value is NA. Most values are NA, because for most event_ids that particular action was not performed (the Pets + Animals feed was not clicked). This will be a sparse dataset.

## Remaining to-dos to get into desired format
1. Separate grouped_logs column on semi-colon - use library(splitstackshape)? check: does each event_id have the same number of logs?
2. Create a column for each key, then add the values to that column
3. Reduce size of sample dataset

## Fixed issues
* Remove semi-colon from end of grouped-logs
* Export flattened data over a wider date range
